---
title: "Vis I"
author: "Jinghan Zhao"
date: "2024-09-26"
output: github_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(ggridges)
```

Import weather data

```{r}
weather_df = 
  rnoaa::meteo_pull_monitors(
    c("USW00094728", "USW00022534", "USS0023B17S"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2021-01-01",
    date_max = "2022-12-31") |>
  mutate(
    name = case_match(
      id, 
      "USW00094728" ~ "CentralPark_NY", 
      "USW00022534" ~ "Molokai_HI",
      "USS0023B17S" ~ "Waterhole_WA"),
    tmin = tmin / 10,
    tmax = tmax / 10) |>
  select(name, id, everything())
```

## Making our first plot

```{r}
ggplot(weather_df, aes(x = tmin, y = tmax)) +
  geom_point()
```

```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax)) +
  geom_point()
```

Save the plot

```{r}
ggp_weather_scatterplot = 
  weather_df %>% 
  ggplot(aes(x = tmin, y = tmax)) +
  geom_point()

ggp_weather_scatterplot
```

Check NAs.

```{r}
weather_df %>% 
  filter(is.na(tmax))
```

## Fancier scatterplots!

`se = FALSE`: not calculate standard error (hardly used)

```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point(alpha = .3, size = .8) +
  geom_smooth(se = FALSE)
```

Where you define aesthetics can matter

(Only define in `geom_point`, but no to do with `geom_smooth`)

```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax)) +
  geom_point(aes(color = name), alpha = .3, size = .8) +
  geom_smooth(se = FALSE)
```

Use faceting real quick

Divide into three plots:

`facet_grid(. ~ name)` instead of `facet_grid(name ~ .)` to plot along columns.

```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point(alpha = .3) +
  geom_smooth(se = FALSE) +
  facet_grid(. ~ name)

weather_df %>% 
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point(alpha = .3) +
  geom_smooth(se = FALSE) +
  facet_grid(name ~ .)

```

Let's make a somewhat more interesting scatterplot

```{r}
weather_df %>% 
  ggplot(aes(x = date, y = tmax, color = name, size = prcp)) +
  geom_point(alpha = .3) +
  geom_smooth(se = FALSE)+
  facet_grid(. ~ name)
```

Learning assessment plots

```{r}
weather_df %>% 
  filter(name == "CentralPark_NY") %>% 
  mutate(
    tmax_fahr = tmax * (9 / 5) + 32,
    tmin_fahr = tmin * (9 / 5) + 32
  ) %>% 
  ggplot(aes(x = tmin_fahr, y = tmax_fahr)) +
  geom_point() +
  geom_smooth(metho = "lm", se = FALSE)
  
```

## Small things

```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax)) +
  geom_point(aes(color = name), alpha = .3, size = .8) +
  geom_smooth(se = FALSE)
```

Look at the density

```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax)) +
  geom_hex()
```

If it's not a variable, use `"..."`. If it's a variable, write directly.

It will define a new variable named `blue` in the second way

```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax)) +
  geom_point(color = "blue")

weather_df %>% 
  ggplot(aes(x = tmin, y = tmax, color = "blue")) +
  geom_point()
```

## Univariate plots

```{r}
weather_df %>% 
  ggplot(aes(x = tmin)) +
  geom_histogram()
```

`fill = name` and `color = name`: color of fill or line

Not overlapping with each other: `position = "dodge"` (but still hardly used)

```{r}
weather_df %>% 
  ggplot(aes(x = tmin, fill = name)) +
  geom_histogram(position = "dodge")
```

How would I fix this (making a better visual comparison) ? maybe facet?

```{r}
weather_df %>% 
  ggplot(aes(x = tmin, fill = name)) +
  geom_histogram() +
  facet_grid(. ~ name)
```

Density plot

```{r}
weather_df %>% 
  ggplot(aes(x = tmin, fill = name)) +
  geom_density(alpha = .3)
```

Box plot

(Colors is not neccessary in this plot because already have `name` in x axis)

```{r}
weather_df %>% 
  ggplot(aes(x = name, y = tmin, fill = name)) +
  geom_boxplot()
```

Violin plot

```{r}
weather_df %>% 
  ggplot(aes(x = name, y = tmin, fill = name)) +
  geom_violin()
```

Ridge plot

```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = name)) +
  geom_density_ridges()
```

Learning assessment univariate plots

(Find a better way...)

```{r}
weather_df %>% 
  ggplot(aes(x = prcp, fill = name)) +
  geom_density(alpha = .3)

weather_df %>% 
  ggplot(aes(x = name, y = prcp)) +
  geom_boxplot()

weather_df %>% 
  filter(prcp > 10, prcp < 1000) %>% 
  ggplot(aes(x = prcp, fill = name)) +
  geom_density(alpha = .3)
```



### Saving and embedding plots

Saving plots

If not define which plot to save, `ggsave()` will save the most recent plot you made.

Instead, define the plot variable to clarify.

```{r}
ggp_weather = 
weather_df %>% 
  ggplot(aes(x = date, y = tmax, color = name)) +
  geom_point()

ggsave("ggp_weather.pdf", ggp_weather, width = 8, height = 6)
```

Embedding plots

To change the picture displayed in R markdown.

`fig.width = 6, fig.asp = .6`: height is 60% of width

```{r, fig.width = 12, fig.height = 12}
weather_df %>% 
  ggplot(aes(x = date, y = tmax, color = name)) +
  geom_point()
```

