---
title: "Vis II"
author: "Jinghan Zhao"
date: "2024-10-01"
output: github_document
---

```{r setup, include = FALSE}
library(tidyverse)
library(patchwork)
library(haven)
```

Import weather data

```{r, message = FALSE}
weather_df = 
  rnoaa::meteo_pull_monitors(
    c("USW00094728", "USW00022534", "USS0023B17S"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2021-01-01",
    date_max = "2022-12-31") |>
  mutate(
    name = case_match(
      id, 
      "USW00094728" ~ "CentralPark_NY", 
      "USW00022534" ~ "Molokai_HI",
      "USS0023B17S" ~ "Waterhole_WA"),
    tmin = tmin / 10,
    tmax = tmax / 10) |>
  select(name, id, everything())
```

## Labels

Make a scatterplot but fancy this time.

```{r}
weather_df %>%
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point(alpha = .3) +
  labs(
    title = "Temperature scatterplot",
    x = "Minimum Temp (C)",
    y = "Maximum Temp (C)",
    color = "Location",
    caption = "Weather data taken from rnoaa package for three stations."
  )
```

## Scales

-- start with `x` and `y`

```{r}
weather_df %>%
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point(alpha = .3) +
  labs(
    title = "Temperature scatterplot",
    x = "Minimum Temp (C)",
    y = "Maximum Temp (C)",
    color = "Location",
    caption = "Weather data taken from rnoaa package for three stations."
  ) +
  scale_x_continuous(
    breaks = c(-15, 0, 20),
    labels = c("-15C", "0", "20C")
  ) +
  scale_y_continuous(
    limits = c(0, 30),
    transform = "sqrt"
  )
```

Look at color:

Use `viridis` package to choose color scheme

```{r}
ggp_scatterplot = 
weather_df %>%
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point(alpha = .3) +
  labs(
    title = "Temperature scatterplot",
    x = "Minimum Temp (C)",
    y = "Maximum Temp (C)",
    color = "Location",
    caption = "Weather data taken from rnoaa package for three stations."
  ) +
  viridis::scale_color_viridis(discrete = TRUE)
```

## Themes

Themes are used to modify non-data elements of a plot.

```{r}
ggp_scatterplot +
  theme(legend.position = "bottom")
```

```{r}
ggp_scatterplot +
  theme_bw() +
  theme(legend.position = "bottom")
```

Order matters...

`theme()` changes a particular element of the plot’s current “theme”. If you call theme to change the some element and then `theme_bw()`, the changes introduced by `theme()` are overwritten by `theme_bw()`.

```{r}
ggp_scatterplot +
  theme(legend.position = "bottom") +
  theme_bw()
```

```{r}
ggp_scatterplot +
  theme_minimal() +
  theme(legend.position = "bottom")
```

```{r}
ggp_scatterplot +
  theme_classic() +
  theme(legend.position = "bottom")
```

```{r}
ggp_scatterplot +
  ggthemes::theme_excel() +
  theme(legend.position = "bottom")
```

## Learning assessment

```{r}
weather_df %>% 
  ggplot(aes(x = date, y = tmax, color = name)) +
  geom_point(alpha = .3) +
  geom_smooth(se = FALSE) +
  viridis::scale_color_viridis(discrete = TRUE) +
  labs(
    x = "Date",
    y = "Maximum Temp (C)",
    title = "Seasonal variation in Max Temp"
  ) +
  theme_minimal() +
  theme(legend.position = "bottom")
```


## Different `geom_*`

Extra bonus stuff in `ggplot`

Use different datasets in different `geom`s

```{r}
central_park_df = 
  weather_df %>% 
  filter(name == "CentralPark_NY")

molokai_df = 
  weather_df %>% 
  filter(name == "Molokai_HI")

molokai_df %>% 
  ggplot(aes(x = date, y = tmax, color = name)) +
  geom_point() +
  geom_line(data = central_park_df)
```



## Multiple panels

```{r}
weather_df %>% 
  ggplot(aes(x = tmax, fill = name)) +
  geom_density() +
  facet_grid(. ~ name)
```

Use `library(patchwork)`

```{r}
ggp_tmax_tmin = 
  weather_df %>% 
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point(alpha = .3)

ggp_tmax_density = 
  weather_df %>% 
  ggplot(aes(x = tmax, fill = name)) +
  geom_density(alpha = .3)

ggp_tmax_date = 
  weather_df %>% 
  ggplot(aes(x = date, y = tmax, color = name)) +
  geom_point() +
  geom_smooth(se = FALSE)

(ggp_tmax_tmin + ggp_tmax_density) / ggp_tmax_date
```

## Setting options

In addition to figure sizing, I include a few other figure preferences in global options declared at the outset of each .Rmd file (this code chunk just gets copy-and-pasted to the beginning of every new file).

```{r, eval = FALSE}
library(tidyverse)

knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```


## Data manipulation

Reorder the factor variables (there are more ways to do it automatically...)

```{r}
weather_df %>% 
  mutate(name = fct_relevel(name, c("Molokai_HI", "CentralPark_NY", "Waterhole_WA"))) %>% 
  ggplot(aes(x = name, y = tmax, fill = name)) +
  geom_violin(alpha = .5)
```


PULSE data next.

```{r}
pulse_df = 
  read_sas("data/public_pulse_data.sas7bdat") %>% 
  janitor::clean_names() %>% 
  pivot_longer(
    cols = bdi_score_bl:bdi_score_12m,
    names_to = "visit",
    values_to = "bdi_score",
    names_prefix = "bdi_score_"
  ) %>% 
  mutate(visit = ifelse(visit == "bl", "00m", visit))

pulse_df %>% 
  ggplot(aes(x = visit, y = bdi_score)) +
  geom_boxplot()
```

Make an FAS plot.

```{r}
litters_df = 
  read_csv("data/FAS_litters.csv", na = c("NA", ".", "")) %>% 
  janitor::clean_names() %>% 
  separate(group, into = c("dose", "tx_day"), 3)

pups_df = 
  read_csv("data/FAS_pups.csv", na = c("NA", ".", "")) %>% 
  janitor::clean_names() %>% 
  pivot_longer(
    pd_ears:pd_walk,
    names_to = "outcome",
    values_to = "pn_day",
    names_prefix = "pd_"
  )

fas_df = 
  left_join(pups_df, litters_df, by = "litter_number")

fas_df %>% 
  drop_na(tx_day) %>% 
  ggplot(aes(x = dose, y = pn_day)) +
  geom_boxplot() +
  facet_grid(tx_day ~ outcome)
```


